Continuous Deductive Reasoning over Streamed RDF Data

### 2015-11-24 article.tex
sooooomany updates

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amscd}

@@

To this end, we say that an interpretation $\mathcal{I}^T$ satisfies a
fact instance $\langle f, t \rangle$ iff
$\langle f, t \rangle \in \mathcal{I}^T$, as in positive Datalog.
We also say that the interpretation $\mathcal{I}^T$ satisfies a rule
$r = \alpha_0 \Leftarrow \alpha_1 \land \ldots \land \alpha_n$ iff, for
every \emph{extended variable substitution} $\bar{s}$ by which
$\{\langle \bar{s}\left( \alpha_1 \right), t_1 \rangle, \dots, \langle \bar{s}\left( \alpha_n \right), t_n \rangle\} \subseteq \mathcal{I}^T$, there exists a fact instance
$\langle \bar{s}\left( \alpha_0 \right), max\left(t_1, \dots, t_n \right) \rangle \in \mathcal{I}^T$.
The meaning of the restriction on the instant associated with the instance of
$\bar{s}\left( \alpha_0 \right)$ in $\mathcal{I}^T$ is better described in terms
of an entailment and its time-annotated justification, as per
\refdef{continuous datalog: instance justification}.

@@

@@

We initially take a {\em Continuous Datalog program} (henceforth a
C-Datalog program) $\Pi^C = \langle R, F^0, S \rangle$ to be a tuple
which comprises a finite set of rules $R$, a finite set of persistent/\emph{static}
fact instances $F^0$ (all of which have the instant $0$, meaning that the facts
instantiated are recognised to be true from the beginning of time and cannot
appear in any state of any window, by \refdef{continuous datalog: window} and
Axioms~\ref{axiom:continuous datalog: window range leq 0} and
\ref{axiom:continuous datalog: window start leq 0}) and an in-order stream
of fact instances $S$.  It should be noted that, in practice, data may be
drawn from more than one source, and so $S$ may be a union of $n$ in-order
streams $S_1,\dots,S_n$ without altering the semantics presented here.

The model theory for a such a C-Datalog program broadly follows that
for a TA-Datalog program; we again take the Herbrand base of all fact
instances $\hat{H}$. Disregarding the order of fact instances in $S$,
a fact instance $\langle e, t \rangle$ is entailed by a program $\Pi^C$
iff each interpretation satisfying every clause in $\Pi^C$ also
satisfies $\langle e, t \rangle$. However, we take the {\em entailment
of the C-Datalog program} $\Pi^C$ to be an in-order stream $E^S$ rather
than a set of fact instances. Given that the instant associated with an
entailed fact instance is the maximum of those associated with
any part of its justification (as in
\refdef{continuous datalog: instance justification}), and $E^S$ is an
in-order stream, $E^S$ expresses the times at which each entailed fact
instance was first justified by $\Pi^C$.

However, this definition of a C-Datalog program does not take into
account any notion of windowing over the stream $S$, and so any fact of
which an instance occurs in $S$ will be considered true from the instant
of its occurrence in $S$.  While this is technically acceptable in the
case of sets of rules without constraints, if a constraint exists and its
body becomes satisfied by some $\Pi^C$ at some instant $i$ then the
program will be unsatisfiable from that instant on, making $E^S$ empty
from then on.  As such, we define a {\em Windowed Continuous Datalog program}
$\Pi^{C, r} = \langle R, F^0, \{ W^{S_1, r_1},\dots, W^{S_n, r_n} \rangle\} \rangle$
(where $r = max(r_1,\dots,r_n)$) to be a tuple with components $R$ and $F^0$ as defined above for a
C-Datalog program, but with each part of $S = S_1 \cup \dots \cup S_n$
associated with the range $r_x \in \mathbb{N}$ of a sliding window over
$S_x$, for $x \in [1,n]$. This specifies that the fact instances
from $S_x$ are to contribute to the entailments of $\Pi^{C,r}$ from the instant
they occur in $S_x$ for a duration of $r_x$ consecutive instants.

@@

From the definition of entailment for TA-Datalog in \refdef{continuous
  datalog: ta datalog program} and the definition of windows in
\refdef{continuous datalog: window}, we can derive the following:

@@

We can reduce a WC-Datalog program $\Pi^{C,r} = \langle R, F^0,
\{ W^{S_1, r_1},\dots, W^{S_n, r_n} \} \rangle$
at an instant $i \in \mathbb{T}$ to its TA-Datalog equivalent by
applying the rules $R$ to the union of the persistent facts $F^0$ and all
finite states $W^{S_x,r_x}_i$ of windows $W^{S_x,r_x}$ over streams $S_x$ of range
$r_x \in \mathbb{N}$, for all $x \in [1,n]$; we refer to this as the {\em instantaneous state} of
the WC-Datalog program, written
$\Pi^{C,r}_i = \langle R, F^0 \cup \mathop{\cup}\limits_{x=1}^n W^{S_j,r_j}_i \rangle$.
As previously shown in \refdef{continuous datalog: ta datalog program}, this may then be further
reduced to the Datalog program
$\Pi = \langle R, strip(F^0 \cup \mathop{\cup}\limits_{x=1}^n W^{S_j,r_j}_i \rangle$.

The instantaneous entailment of a $\Pi^{C,r}_i$ may be divided into two
semantically distinct portions: the \emph{persistent} portion is the
entailment of ${\langle R, F^0 \rangle \models E^0}$, as in a WC-Datalog
program, which is guaranteed to be part of the entailment of every consistent
state due to the monotonicity of positive Datalog; the \emph{transient} portion
is the set $E^T_{i}$ of entailed fact instances whose justifications are at
least partially contained in $\mathop{\cup}\limits_{x=1}^n W^{S_j,r_j}_i$.
{\nobreak We define $E^T_i$ as:
\[
E^T_i = \left\{ e \,\middle\vert\, \Pi^{C,r}_i \models e \land \left( J^T \cap \bigcup_{x=1}^n W^{S_j,r_j}_i \neq \varnothing \right) \right\}
\]
where $J^T$ is a justification for the entailment of $e$ by
$\Pi^{C,r}_i$.} By this definition and Axiom~\ref{axiom:continuous
  datalog: entailment disjointness}, $E^0$ is guaranteed to be
disjoint from every $E^T_i$ for every state $\Pi^{C,r}_i$ of every
WC-Datalog program $\Pi^{C,r}$.

@@

Thus far the translation from instantaneous sets of entailments to
streams of entailments has been stated axiomatically as the definition
of those streams.
However, it follows from these axioms that the instantaneous set
$E^T_{i}$ of instances entailed at any given instant $i$ by any single-window program
$\Pi^{C,r} = \langle R, F^0, \{ \langle S, r \rangle \} \rangle$ may be reconstructed from
windows of sufficient range {\nobreak over its entailed streams $P$ and $N$ at
$i$:
\[ E^T_{i} = W^{P,x}_{i} \setminus W^{N,y}_{i} \]
where ${y \geqslant x \geqslant r}$}. The minimum range $x$ is such to achieve
the \emph{complete} set of entailment instances $E^T_{i}$ of the
$\Pi^{C,r}_{i}$. In addition, the minimum range $y$ is such to achieve the
\emph{sound} set $E^T_{i}$.  It should be noted that the proof holds in the
case of $n$ windows of varying range, with $x \geqslant max(r_0,\dots,r_n)$.
We capture this derivation in the form of a \emph{valid window}, as per
\refdef{continuous datalog: valid window}.

@@

The decomposition of any stream into any number of more
sparse streams without loss of information allows us to define the
state of a valid window $W^{P \setminus N}$ as a union of the states
of sliding windows over each of the singleton streams
$(\langle f, t \rangle \mapsto t)$, the union of which is equal to $P$,
where $P(\langle f, t \rangle) = t$ means that $P$ maps
$\langle f, t \rangle$ to $t$.  The range of the sliding
window over each singleton stream is the difference between
the instant at which $\langle f, t \rangle$ occurs in $P$ ($t$, as $P$
is in-order) and that at which it occurs in $N$:
\[
  W^{P \setminus N}_i = \left\{ \langle f, t \rangle \,\middle\vert\, \begin{split}
    P(\langle f, t \rangle) = t \land \\
    \langle f, t \rangle \in W^{\left( \langle f, t \rangle \mapsto t \right),N(\langle f, t \rangle) - t}_i
  \end{split}\right\}
\]
Treated this way, the entailment of one WC-Datalog program, e.g. that in
\refeqn{continuous datalog: example WC-Datalog program},
may then be taken as a part or all of the data in a successive WC-Datalog
program, without modification of the existing semantics or loss of information
regarding the validity of the entailments.  However, in the interest
of being concise, such a successive WC-Datalog program may express valid
windows alongside any sliding windows. For example, 
a program composed of the set of rules $R^\prime \neq R$, the set of base facts
$F^{0\prime} \neq F^0$, a sliding window of range $x \geqslant r$ over the
in-order stream $S^\prime \neq S$, and the entailment of
$\Pi^{C,r} \models \langle E^0, P, N \rangle$, would be expressed as in \refeqn{continuous datalog: successive WC-Datalog program}.
\begin{gather}
  \labeleqn{continuous datalog: example WC-Datalog program}
    \Pi^{C,r} \langle R, F^0, \{W^{S,r}\} \rangle \models \langle E^0, P, N \rangle \\
  \labeleqn{continuous datalog: successive WC-Datalog program}
    \Pi^{C,x} = \langle R^\prime, F^{0\prime} \cup E^0, \{ W^{P \setminus N}, W^{S^\prime,x} \} \rangle \models \langle E^{0\prime}, P^\prime, N^\prime \rangle
\end{gather}

In conclusion, given a Continuous Datalog program $\Pi^C = \langle R,
F^0, S_0 \cup \dots \cup S_n \rangle$, we can apply to it a set of
windows of range $r_0,\dots,r_n$ to yield a Windowed Continuous Datalog
program $\Pi^{C,r} = \langle R, F^0, \{ W^{S_0, r_0}, \dots,
W^{S_n, r_n} \} \rangle$. We can evaluate these windows at
an instant $i \in \mathbb{T}$ to yield a Time-Annotated Datalog program
$\Pi^T = \langle R, F^0 \cup W^{S_0,r_0}_i \cup \dots \cup W^{S_n,r_n}_i
\rangle$, from we which can then strip the time annotations to yield a
conventional Datalog program $\Pi = \langle R, F \rangle$. The
relationship between these $\ast$-Datalog programs and their
entailments is expressed in the commutative diagram in
\reffig{CD-TD-D}.

@@

\begin{nestedsection}{CoDeR Model for Continuous Deductive Reasoning}{model}
  CoDeR is an abstract model for continuous deductive reasoning over streamed RDF data that expresses the semantics of Continuous Datalog, thus supporting continuous reasoning expressed in both RIF-Core and OWL 2 RL.
  It is composed of a stream-based data model for expressing the streams $S$, $P$ and $N$ of each ${\Pi^{C,r} = \langle R,F^0,S,r \rangle \models \langle E^0,P,N \rangle}$, and a minimal algebra of stream-to-stream operators for expressing the persistent set of rules $R$.
  It casts the problem of applying those rules to streamed RDF data as that of iterative pattern matching, in a manner inspired by the Rete pattern matching algorithm \citep{forgy79}, though it is meant to support any such algorithm rather than prescribing one itself.

  % Stream-based Data Model
  \begin{nestedsection}{Data Model}{model: data}
    The data model of CoDeR is composed of multiple-source multiple-sink streams.
    These streams are divided into two classes: \emph{input}/\emph{output} streams and \emph{inter-operator} streams.

    In the first case, input/output streams provide the semantics of the streams $S$, $P$ and $N$ for any $\Pi^{C,r}$ expressed by a CoDeR-based system.
    These streams are sequences of instances of heterogeneous RDF graphs compatible with the ongoing work on query semantics by the RSP working group\footnote{\url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}};
    an instance of an RDF graph is an RDF graph annotated with a time interval ${(e,n]}$: ${\langle \{(s,p,o),\dots,(s',p',o')\},e,n \rangle}$.
    As the atomic units of Continuous Datalog are instances of facts, being triples in RDF, graph instances arriving on the input stream must be translated into sequences of fact instances using the graph-to-triples operator described in \refsec{model: algebra}, where each fact instance inherits the time interval of the graph instance in which it originated.

    The interval ${(e,n]}$ for any graph or triple instance is the the period of time for which the data is \emph{valid} \citep{SemanticStreamingManagement,sparkwave} according to the ${\Pi^{C,r} = \langle R, F^0, S, r \rangle}$ expressed.
    For graphs on the input stream, $e$ is the instant at which a given graph was asserted, and ${n = e + r}$ is the instant at which it will leave the semantic sliding window over $S$ of $\Pi^{C,r}$ (or become ``negated'').
    For graphs on the output stream, $e$ is the latest instant at which any part of the justification of a given graph was asserted, and $n$ is the earliest instant at which any part of its justification will leave the sliding window.
    \reffig{intersected-intervals} provides an example of this for some ${\Pi^{C,r} = \langle \{c \Leftarrow a \land b\},\{\},[\langle a,1 \rangle,\langle b,2 \rangle ], 3\rangle}$ over instants ${i = 1 \dots 4}$.
    \begin{figure}
      \centering
      \includegraphics[width=0.3\textwidth]{intersected-intervals}
      \caption{Derivation of valid-time for entailed fact instance from those of its justification.}
      \labelfig{intersected-intervals}
    \end{figure}
    \begin{align*}
      \Pi^{C,r}_{1} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle \} \rangle & \models & \{ \} \\
      \Pi^{C,r}_{2} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
      \Pi^{C,r}_{3} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
      \Pi^{C,r}_{4} & = \langle \{c \Leftarrow a \land b\}, \{ \langle b, 2 \rangle \} & \models & \{ \}
    \end{align*}
    It is clear that a CoDeR output stream expresses the semantic stream $P$ entailed by some $\Pi^{C,r}$.
    However, as the semantic stream $N$ is defined as the sequence of entailments of some $\Pi^{C,r}$ ordered by the instant $n$ at which $W^{S,r}_{n}$ \emph{ceased} to justify them, the output stream of a CoDeR-based system also expresses the stream $N$.
    As such, the annotation of streamed RDF graphs with the interval $(e,n]$ gives the streams of CoDeR a dual semantics with regard to Continuous Datalog.

    In the second case, inter-operator streams are produced and consumed by the stream-to-stream operators of CoDeR's minimal algebra.
    These streams are sequences of the intermediate results of processing between steps in the production of entailments.
    As the production of entailments in CoDeR is cast as an iterative pattern matching problem, all intermediate results of this process will also take the form of instances of RDF graphs;
    an inter-operator stream is a sequence of instances of homogeneous RDF graphs, each matching the same graph pattern, that being a sub-pattern of the body of one of the rules $R$ of the expressed $\Pi^{C,r}$.
    These instances of RDF graphs could be considered to be ``entailed'' by the sub-pattern of the stream on which they appear and the presence of valid instances of their constituent RDF triples within the system;
    as such, the same dual semantics of valid-times may be applied to these intermediate matches as to the true entailments of the model found on output streams.
    It follows that inter-operator streams are functionally the same as input/output streams, only being distinct in their semantics:
    the former are sequences of homogeneous RDF graphs that hold no value outside of a system, while the latter contain heterogeneous graphs that are either the entailments of the system or the assertions of its predecessor in the stream pipeline.

>>>>

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amscd}
\usepackage{mathtools}

@@

To this end, we say that an interpretation $\mathcal{I}^T$ satisfies a
fact instance $\langle f, t \rangle$ iff
$\langle f, t \rangle \in \mathcal{I}^T$, as in positive Datalog.
We also say that the interpretation $\mathcal{I}^T$ satisfies a rule
$r = \alpha_0 \Leftarrow \alpha_1 \land \ldots \land \alpha_n$ iff, for
every \emph{extended variable substitution} $\bar{s}$ by which
$\{\langle \bar{s}\left( \alpha_1 \right), t_1 \rangle, \dots, \langle \bar{s}\left( \alpha_n \right), t_n \rangle\} \subseteq \mathcal{I}^T$, there exists a fact instance
$\langle \bar{s}\left( \alpha_0 \right), max\left(t_1, \dots, t_n \right) \rangle \in \mathcal{I}^T$.
The meaning of the restriction on the instant associated with the instance of
$\bar{s}\left( \alpha_0 \right)$ in $\mathcal{I}^T$ is better described in terms
of the time-annotated justification of an entailed fact instance, as per
\refdef{continuous datalog: instance justification}.

@@

From the definition of entailment for TA-Datalog in \refdef{continuous
  datalog: ta datalog program} and the definition of the states of sliding windows in
\refdef{continuous datalog: window}, we can derive the following:

@@

We initially take a {\em Continuous Datalog program} (henceforth a
C-Datalog program) $\Pi^C = \langle R, F^0, S \rangle$ to be a tuple
which comprises a finite set of rules $R$, a finite set of persistent/\emph{static}
fact instances $F^0$ (all of which have the instant $0$, meaning that the facts
instantiated are recognised to be true from the beginning of time and cannot
appear in any state of any window, by \refdef{continuous datalog: window} and
Axioms~\ref{axiom:continuous datalog: window range leq 0} and
\ref{axiom:continuous datalog: window start leq 0}) and an in-order stream
of fact instances $S$.  It should be noted that, in practice, data may be
drawn from more than one source, and so $S$ may be a union of $s$ in-order
streams $S_1,\dots,S_s$ without altering the semantics presented here.

The model theory for a such a C-Datalog program broadly follows that
for a TA-Datalog program; we again take the Herbrand base of all fact
instances $\hat{H}$. Disregarding the order of fact instances in $S$,
a fact instance $\langle e, t \rangle$ is entailed by a program $\Pi^C$
iff each interpretation satisfying every clause in $\Pi^C$ also
satisfies $\langle e, t \rangle$. However, we take the {\em entailment
of the C-Datalog program} $\Pi^C$ to be an in-order stream $E^S$ rather
than a set of fact instances. Given that the instant associated with an
entailed fact instance is the maximum of those associated with
any part of its justification (as in
\refdef{continuous datalog: instance justification}), and $E^S$ is an
in-order stream, $E^S$ expresses the times at and order in which each
entailed fact instance was first justified by $\Pi^C$.

However, this definition of a C-Datalog program does not take into
account any notion of windowing over the stream $S$, and so any fact of
which an instance occurs in $S$ will be considered true from the instant
of its occurrence in $S$.  While this is technically acceptable in the
case of sets of rules without constraints, if a constraint exists and its
body becomes satisfied by some $\Pi^C$ at some instant $i$ then the
program will be unsatisfiable from that instant on, making $E^S$ empty
from then on.  As such, we define a {\em Windowed Continuous Datalog program}
$\Pi^{C, r} = \langle R, F^0, \{ W^{S_x, r_x} \mid x \in (1,s) \} \rangle\} \rangle$
(where $r = max(r_1,\dots,r_s)$) to be a tuple with components $R$ and $F^0$ as defined above for a
C-Datalog program, but with each part of $S = \mathop{\cup}^s_{x=1}S_x$
associated with the range $r_x \in \mathbb{N}$ of a sliding window over
$S_x$, for $x \in (1,s)$. This specifies that the fact instances
from $S_x$ are to contribute to the entailments of $\Pi^{C,r}$ from the instant
they occur in $S_x$ for a duration of $r_x$ consecutive instants.

@@

@@

We can reduce a WC-Datalog program $\Pi^{C,r} = \langle R, F^0,
\{ W^{S_x, r_x} \mid x \in (1,s) \} \rangle$
at an instant $i \in \mathbb{T}$ to its TA-Datalog equivalent by
applying the rules $R$ to the union of the persistent facts $F^0$ and all
finite states $W^{S_x,r_x}_i$ of windows $W^{S_x,r_x}$ over streams $S_x$ of range
$r_x \in \mathbb{N}$, for all $x \in (1,s)$; we refer to this as the {\em instantaneous state} of
the WC-Datalog program, written
$\Pi^{C,r}_i = \langle R, F^0 \cup \mathop{\cup}\limits_{x=1}^s W^{S_x,r_x}_i \rangle$.
As previously shown in \refdef{continuous datalog: ta datalog program}, this may then be further
reduced to the Datalog program
$\Pi = \langle R, strip(F^0 \cup \mathop{\cup}\limits_{x=1}^n W^{S_j,r_j}_i \rangle$.

The instantaneous entailment of a $\Pi^{C,r}_i$ may be divided into two
semantically distinct portions: the \emph{persistent} portion is the
entailment of ${\langle R, F^0 \rangle \models E^0}$, as in a WC-Datalog
program, which is guaranteed to be part of the entailment of every consistent
state due to the monotonicity of positive Datalog; the \emph{transient} portion
is the set $E^T_{i}$ of entailed fact instances whose justifications are at
least partially contained in $\mathop{\cup}\limits_{x=1}^n W^{S_x,r_x}_i$.
{\nobreak We define $E^T_i$ as:
\[
E^T_i = \left\{ e \,\middle\vert\, \Pi^{C,r}_i \models e \land \left( J^T \cap \bigcup_{x=1}^n W^{S_x,r_x}_i \neq \varnothing \right) \right\}
\]
where $J^T$ is a justification for the entailment of $e$ by
$\Pi^{C,r}_i$.} By this definition and Axiom~\ref{axiom:continuous
  datalog: entailment disjointness}, $E^0$ is guaranteed to be
disjoint from every $E^T_i$ for every state $\Pi^{C,r}_i$ of every
WC-Datalog program $\Pi^{C,r}$.

@@

Thus far the translation from instantaneous sets of entailments to
streams of entailments has been stated axiomatically as the definition
of those streams.
However, it follows from these axioms that the instantaneous set
$E^T_{i}$ of instances entailed at any given instant $i$ by any single-window program
$\Pi^{C,r} = \langle R, F^0, \{ \langle S, r \rangle \} \rangle$ may be reconstructed from
windows of sufficient range over its entailed streams $P$ and $N$ at
$i$, as expressed in \refeqn{continuous datalog: entailment equivalence}.
{\nobreak\begin{equation} \labeleqn{continuous datalog: entailment equivalence}
  E^T_{i} = W^{P,x}_{i} \setminus W^{N,y}_{i}
\end{equation}
where ${y \geqslant x \geqslant r}$}. The minimum range $x$ is such to achieve
the \emph{complete} set of entailment instances $E^T_{i}$ of the
$\Pi^{C,r}_{i}$. In addition, the minimum range $y$ is such to achieve the
\emph{sound} set $E^T_{i}$.  It should be noted that the proof holds in the
case of $n$ windows of varying range, with $x \geqslant max(r_1,\dots,r_n)$.
We capture this derivation in the form of a \emph{valid window}, as per
\refdef{continuous datalog: valid window}.

@@

The decomposition of any stream into any number of more
sparse streams without loss of information allows us to alternatively define the
state of the valid window $W^{P \setminus N}$ at any instant $i \in \mathbb{T}$
as the union of the states of the set of sliding windows $\mathcal{W}^{P \setminus N}$
over each of the singleton streams $(\langle f, t \rangle \mapsto t^\prime)$
at $i$, the union of these streams being equal to $P$.  Remembering that
$P(\langle f, t \rangle) = t^\prime$ means that $\langle f, t \rangle$ occurs
in $P$ at $t^\prime$, the range of each sliding window in $\mathcal{W}^{P \setminus N}$
is the difference $d$ between $P(\langle f, t \rangle)$ (which equals $t$, as $P$ is
in-order) and $N(\langle f, t \rangle)$.  This difference is always positive by Axiom
\ref{axiom:continuous datalog: entailment precedes negation}, and bounded at the
upper limit $r$ of the program $\Pi^{C,r} \models \langle E^0, P, N \rangle$ by
\refdef{continuous datalog: CDP}.
\begin{gather}
  W^{P \setminus N}_i = \bigcup_{W^{p,d} \in \mathcal{W}^{P \setminus N}} W^{p,d}_i \\
  \mathcal{W}^{P \setminus N} = \left\{ W^{p,d} \,\middle\vert\, \begin{split}
    \langle f, t \rangle \in P \land \\
    p = (\langle f, t \rangle \mapsto t) \land \\
    d = N(\langle f, t \rangle) - t
  \end{split}\right\}
\end{gather}
Treated this way, the entailment of one WC-Datalog program, e.g. that in
\refeqn{continuous datalog: example WC-Datalog program}, may then be taken as a
part or all of the data in a successive WC-Datalog program, without modification
of the existing semantics or loss of information regarding the validity of the
entailments.
% However, in the interest
% of being concise, such a successive WC-Datalog program may express valid
% windows alongside any sliding windows.
For example, 
a program composed of the set of rules $R^\prime$, the set of base facts
$F^{0\prime}$, a sliding window of range $x \geqslant r$ over the
in-order stream $S^\prime$, and the entailment of
$\Pi^{C,r} \models \langle E^0, P, N \rangle$, would be expressed as in \refeqn{continuous datalog: successive WC-Datalog program}.
\begin{gather}
  \labeleqn{continuous datalog: example WC-Datalog program}
    \Pi^{C,r} \langle R, F^0, \{W^{S,r}\} \rangle \models \langle E^0, P, N \rangle \\
  \labeleqn{continuous datalog: successive WC-Datalog program}
    \Pi^{C,x} = \langle R^\prime, F^{0\prime} \cup E^0, \{ W^{S^\prime,x} \} \cup \mathcal{W}^{P \setminus N} \rangle \models \langle E^{0\prime}, P^\prime, N^\prime \rangle
\end{gather}

In conclusion, given a Continuous Datalog program $\Pi^C = \langle R,
F^0, S_1 \cup \dots \cup S_n \rangle$, we can apply to it a set of
windows of range $r_1,\dots,r_n$ to yield a Windowed Continuous Datalog
program $\Pi^{C,r} = \langle R, F^0, \{ W^{S_1, r_1}, \dots,
W^{S_n, r_n} \} \rangle$. We can evaluate these windows and a valid window
over the streamed entailment at an instant $i \in \mathbb{T}$ to yield a
Time-Annotated Datalog program $\Pi^T = \langle R, F^0 \cup W^{S_1,r_1}_i
\cup \dots \cup W^{S_n,r_n}_i \rangle$.  From this we can then strip the
time annotations to yield a conventional Datalog program $\Pi = \langle
R, F \rangle$, where $F = strip(F^0 \cup W^{S_1,r_1}_i \cup \dots \cup
W^{S_n,r_n}_i)$. The relationship between these $\ast$-Datalog programs,
mirrored in the relationship between their entailments, is expressed in
the commutative diagram in \reffig{CD-TD-D}.

\begin{nestedsection}{In the Semantic Web: Windowed Continuous RIF-Core}{continuous datalog: rif}
  It has been noted that RIF-Core \citep{w3crifcore} shares the semantics
  of Positive Datalog, extended solely with a set of built-in predicates
  describing the persistent characteristics of and relationships between
  all literal values. These predicates require an updated notion of rule
  \emph{safety}, but this is only relevant to the formulation of rules,
  at the semantic level; as this is unchanged in any $\ast$-Datalog from
  that of positive Datalog, the extended form of rule safety does not
  impact the semantics of satisfiability and entailment of $\ast$-RIF-Core
  programs, as an extension of $\ast$-Datalog programs.

  With regards to the notation of WC-Datalog, these built-in predicates
  may be represented by the set of facts $B$, or the set of persistent
  fact instances ${B^0 = \{ \langle b, 0 \rangle \mid b \in B \}}$, that
  are the collected extension of those predicates. A RIF-Core program may
  therefore be represented as a Datalog program
  ${\Pi = \langle R, F \cup B \rangle}$, a Time-Annotated RIF-Core program
  as a TA-Datalog program ${\Pi^{T} = \langle R, F^T \cup B^0 \rangle}$, a
  Continuous RIF-Core program as a C-Datalog program
  ${\Pi^{C} = \langle R, F^0 \cup B^0, S_1 \cup \dots \cup S_n \rangle}$
  and a Windowed Continuous RIF-Core program as a WC-Datalog program
  ${\Pi^{C,r} = \langle R, F^0 \cup B^0, \{W^{S_1,r_1}, \dots, W^{S_n,r_n}\} \rangle}$.
  Furthermore, ${\langle R, B \rangle \models \varnothing}$ and
  ${\langle R, B^0 \rangle \models \varnothing}$, for any set $R$ of
  \emph{safe} RIF-Core rules. As such, the form of entailment for each
  type of RIF-Core program is the same as that for the equivalent variety
  of Datalog program:
  \begin{align*}
    \Pi =&& \langle R, F \cup B \rangle &\models E \\
    \Pi^{T} =&& \langle R, F^T \cup B^0 \rangle &\models E^T \\
    \Pi^{C} =&& \langle R, F^0 \cup B^0, S \rangle &\models E^S \\
    \Pi^{C,r} =&& \langle R, F^0 \cup B^0, S, r \rangle &\models \langle E^0, P, N \rangle \\
    \Pi^{C,r}_i =&& \langle R, F^0 \cup B^0 \cup W^{S,r}_i \rangle &\models E^0 \cup E^T_i
  \end{align*}

  The reason for maintaining the distinction between $F$/$F^T$/$F^0$ and
  $B$/$B^0$ is the fact that the sets $F$/$F^T$/$F^0$ are strictly finite,
  whereas many of the extensions of the built-in predicates in $B$/$B^0$ are
  strictly infinite. As such, these predicates are not expected to be
  extended in any real collection of data, and so propositions based on
  built-in predicates cannot be satisfied through an inspection of collected
  facts, as is the case for both the facts in any $F$ and fact instances in
  any $F^T$, $F^0$ and $S$; instead, propositions that extend built-in
  predicates can be satisfied by a given interpretation by performing
  calculations on arguments substituted by a variable substitution $\bar{s}$,
  where the propositions in the same rule body that do not extend built-in
  predicates are satisfied by the interpretation and the substitution
  $\bar{s}$. This is the reason that propositions that extend built-in
  predicates do not contribute to variable safety in RIF-Core, in the majority
  of cases.

  As can be seen, the built-in predicates of RIF-Core do not alter the
  semantics of entailment in positive Datalog, nor do they alter these
  semantics in its time-annotated and continuous extensions here presented.
  Rather, they provide an additional set of persistent facts that may be
  utilised within the set $R$ of rules for any *-Datalog program.
\end{nestedsection}

@@

\begin{nestedsection}{CoDeR Model for Continuous Deductive Reasoning}{model}
  \emph{CoDeR} is an abstract model for \emph{Co}ntinuous \emph{De}ductive \emph{R}easoning over streamed RDF data that expresses the semantics of Windowed Continuous RIF-Core, thus supporting continuous reasoning with general knowledge expressed in both RIF-Core and OWL 2 RL.
  It is composed of a stream-based semantic data model and a minimal algebra of stream-to-stream operators, the former for expressing the persistent fact instances $F^0$ and $E^0$, the streams $S_i$, $P$ and $N$ and the window ranges $r_i$ for $i \in (1,s)$ of each ${\Pi^{C,r} = \langle R,F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in (1,s) \} \rangle \models \langle E^0,P,N \rangle}$, the latter for expressing the persistent set of rules $R$, the built-in predicates of RIF-Core $B^0$ and enforcing the window range $r$.
  It casts the problem of applying those rules to streamed RDF data as that of iterative pattern matching, in a manner inspired by the Rete pattern matching algorithm \citep{forgy79}, though it is meant to support any such algorithm rather than prescribing one itself.

  % Stream-based Data Model
  \begin{nestedsection}{Data Model}{model: data}
    The data model of CoDeR is composed of multiple-source multiple-sink data streams, as shown in \reffig{CoDeR: stream}.
  The data model of CoDeR is composed of multiple-source multiple-sink data streams, as shown in \reffig{CoDeR: stream}.
  These streams are based on the notion of in-order streams presented in \refdef{continuous datalog: stream}, but are divided into two classes: \emph{input}/\emph{output} streams and \emph{inter-operator} streams.
  In addition, the notion of an \emph{instance} is extended in \refdef{CoDeR: graph instance} from that of an instance of a \emph{fact} that became true at a specific instant, as presented in \refdef{continuous datalog: fact instance}.
  The streams of CoDeR are streams of this extended form of instance, as per \refdef{CoDeR: stream}, which are compatible with the ongoing work on continuous query semantics by the RSP working group\footnote{According to \url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}, 4th Nov 2015, a CoDeR graph instance may be expressed as an RSP graph with two timestamps predicated as `became valid' and `ceased to be valid'.}.
  \begin{definition}[Graph Instance]\labeldef{CoDeR: graph instance}
    A \emph{graph instance} is an RDF graph with some additional temporal annotation regarding both the instant from which it became valid and the instant at which it will cease to be valid;
    it is a triple composed of the graph ${\{(s,p,o),\dots,(s',p',o')\}}$ and the pair of instants $e$ and $n$ that bound the interval ${(e,n]}$ for which the graph is valid, written ${\langle \{(s,p,o),\dots,(s',p',o')\},e,n \rangle}$.
    The names $e$ and $n$ are so named because they relate to the instants of the instance's {$e$ntailment} and {$n$egation} by the program.
    It should be noted that validity is a stronger concept than truth, as all facts within a graph that is valid in a system are understood to be true, but facts may be true despite not being represented in a graph that is valid within the system.
    Two graph instances containing the same facts and annotation for $e$ and $n$ are considered identical;
    two graph instances containing the same facts but different annotations for $e$ and/or $n$ are considered to be distinct.
  \end{definition}
  \begin{definition}[CoDeR Stream]\labeldef{CoDeR: stream}
    All \emph{streams} of CoDeR are sequences of graph instances, using the definition of \emph{sequence} used in \refdef{continuous datalog: stream} of streams in WC-Datalog.
    They are also \emph{in-order} with regards to their temporal annotation $e$, but potentially \emph{out-of-order} with regards to their temporal annotation $n$.
  \end{definition}
  \begin{definition}[Empty Stream]\labeldef{CoDeR: empty stream}
    The \emph{empty stream} is a CoDeR stream that contains no graph instances, and is denoted by the symbol $\varnothing$.
    This means that the stream will contain no graph instances that map to instant $i$, for all instants ${i \in \mathbb{T}}$.
    This stream may occur in specific expressions of any program, as described in \refsec{CoDeR: transformations}.
  \end{definition}

  It should be noted that the atomic units of WC-RIF-Core are instances of facts, being \emph{triples} in RDF, and so graph instances must be translated into sequences of fact instances in order to truly express the stream $S$ of a program $\Pi^{C,r}$.
  This is achieved by the graph-to-triples operator ${{}_g\pi_t}$ detailed in \refsec{CoDeR: graph-to-triples}, which produces a stream of graph instances in which each instantiated graph contains a single triple, thereby expressing a single fact.

  \begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{logical-stream}
    \caption[CoDeR Data Stream]{A stream of the CoDeR data model, contributed to by two sources and consumed from by two sinks.}
    \labelfig{CoDeR: stream}
  \end{figure}
